### 2024-07-03 paper.tex
Initial commit

\documentclass[sigplan,10pt,anonymous,review]{acmart}
\settopmatter{printfolios=true,printccs=false,printacmref=false}

\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc}
\usepackage[british]{babel}
\usepackage{csquotes}
\usepackage{subcaption}
\usepackage{microtype}
\usepackage{balance}
\usepackage{textgreek}

\bibliographystyle{ACM-Reference-Format}

\hyphenation{meta-va-ri-able meta-va-ri-ables us-abil-i-ty gen-er-al-i-sa-tion
  none-the-less semi-group semi-groups rel-a-tive in-duc-tive rel-e-vant
  tran-si-tive-ly nor-ma-li-sa-tion de-fi-ni-tion de-fi-ni-tions no-ta-bly}

\begin{document}

\begin{abstract}
  The abstract.
\end{abstract}

\title{Proof Script Generation for Aesop}

\author{Jannis Limperg}
\affiliation{%
  \institution{University of Munich (LMU)}
  \department{Department of Computer Science}
  \streetaddress{TODO}
  \city{TODO}
  \postcode{TODO}
  \country{Germany}}
\email{jannis.limperg@lmu.de}
\orcid{0000-0002-8861-5231}

\maketitle

\section{Introduction}

Hello world!

\bibliography{lit}

\end{document}